---
title: "Sample Project A Analyses with dm1 data"
author: "Thomas E. Love"
date: "10/28/2021"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
        number_sections: TRUE
        code_folding: show
---

## Important Note {-}

This is a very quick example. I'm showing the minimum standard, not what we're hoping you will be able to do. But I want to increase the chances that you'll meet the minimum standard for a B on this part of the project.

# Setup {-}

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
```

```{r}
library(broom)
library(tidyverse)

theme_set(theme_bw())
```

## Ingesting the Data

```{r}
dm1 <- read_rds("data/dm1.Rds")
```

## Summarizing the Data

```{r}
summary(dm1)
```
# A Model for a Quantitative Outcome using a Quantitative Predictor

## Initial Visualization

```{r}
ggplot(data = dm1, aes(x = age, y = a1c)) +
    geom_point()
    geom_smooth(method = "lm", formula = y ~ x, 
                se = TRUE, col = "red")
```

## Fitting a Model

```{r}
mod_age <- lm(a1c ~ age, data = dm1)
```

```{r}
tidy(mod_age, conf.int = TRUE, conf.level = 0.90)
```

```{r}
glance(mod_age)
```

```{r}
summary(mod_age)
```

## Residual Plots

```{r}
plot(mod_age, which = c(1:2))
```


# A Model for a Quantitative Outcome using a Multi-Categorical Predictor

## Initial Visualization

```{r}
ggplot(dm1, aes(x = income, y = a1c)) +
    geom_violin() +
    geom_boxplot(width = 0.3)
```

## Fitting a Model

```{r}
mod_inc <- lm(a1c ~ income, data = dm1)
```

```{r}
summary(mod_inc)
```

## ANOVA comparisons

```{r}
anova(mod_inc)
```

## Tukey HSD pairwise comparisons

```{r}
tuk_inc <- TukeyHSD(aov(a1c ~ income, data = dm1), 
                    ordered = TRUE, conf.level = 0.90)

tuk_inc
```

```{r}
plot(tuk_inc)
```

```{r}
dm1 <- dm1 %>%
    mutate(inc2 = fct_recode(income,
                             Low = "Below_30K",
                             Mid = "Between_30-50K",
                             High = "Higher_than_50K"))

dm1 %>% count(income, inc2)
```

```{r}
tuk_inc2 <- TukeyHSD(aov(a1c ~ inc2, data = dm1), 
                    ordered = TRUE, conf.level = 0.90)

plot(tuk_inc2)
```

## Residual Plots

```{r}
par(mfrow = c(1,2))
plot(mod_inc, which = c(1:2))
par(mfrow = c(1,1))
```

# A Model for A1c, using Age, and adjusting for Income

## Initial Visualization

```{r}
ggplot(data = dm1, aes(x = age, y = a1c, col = income, group = income)) +
    geom_point() +
    geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
    facet_wrap(~ income)
```

## Fitting a Model

```{r}
mod_2 <- lm(a1c ~ age * income, data = dm1)
```

```{r}
summary(mod_2)
```

```{r}
tidy(mod_2, conf.int = TRUE, conf.level = 0.90) %>%
    select(estimate, conf.low, conf.high)
```

```{r}
glance(mod_2)
```

```{r}
anova(mod_2)
```

## Making Predictions at the median age, for each income group

```{r}
new_dat <- tibble(age = c(56, 56, 56),
                  income = c("Higher_than_50K", 
                             "Between_30-50K", 
                             "Below_30K"))

new_dat
```

```{r}
augment(mod_2, newdata = new_dat)
```


## Residual Plots

```{r}
par(mfrow = c(1,2))
plot(mod_2, which = c(1:2))
par(mfrow = c(1,1))
```


